A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-02-11 sections/Metatheory.tex
0211 TODOs

The transfer $\Gm\sto\Om$ is not a function.\bruno{Do you mean here
  ``is not deterministic''? Would that be more appropriate?}
On the one hand, there are infinite possibilities to instantiate an existential variable,
thus a judment $\Gm$ could be transferred to different $\Om$'s.
On the other hand, any valid monotype in $\Om$ can be represented by an existential variable in $\Gm$,
thus different $\Gm$'s could be transferred to a single $\Om$.

@@

Figure~\ref{fig:decl:worklist} relates the worklist-style declarative system
to the original declarative system.
A successful reduction means a valid variable declaration,
a valid subtyping or checking judgment, or a successful guessing for inferences.
Additionally, judgment chains are unfolded to match the original declarative rules.
Although $\Om$ is defined differently from $\Psi$,
our declarative worklist $\Om$ contains a ordered collection of variable definitions,
which naturally fits declarative subtyping and typing.
\bruno{I think that to make things precise what you want to show is
  that you can always define a erasure function that drops the
  judgements from the worklist, and retains only the variable
  declarations (thus becomes a declarative environment). Isn't this
  what you do in your formalization?}

\subsection{Soundness}

>>>>

The transfer $\Gm\sto\Om$ is not deterministic.
% \bruno{Do you mean here
%   ``is not deterministic''? Would that be more appropriate?}
On the one hand, there are infinite possibilities to instantiate an existential variable,
thus a judment $\Gm$ could be transferred to different $\Om$'s.
On the other hand, any valid monotype in $\Om$ can be represented by an existential variable in $\Gm$,
thus different $\Gm$'s could be transferred to a single $\Om$.

@@

Figure~\ref{fig:decl:worklist} relates the worklist-style declarative system
to the original declarative system.
A successful reduction means a valid variable declaration,
a valid subtyping or checking judgment, or a successful guessing for inferences.
Additionally, judgment chains are unfolded to match the original declarative rules.
Although $\Om$ is defined differently from $\Psi$,
our declarative worklist $\Om$ contains a ordered collection of variable definitions,
which naturally fits declarative subtyping and typing.
\bruno{I think that to make things precise what you want to show is
  that you can always define a erasure function that drops the
  judgements from the worklist, and retains only the variable
  declarations (thus becomes a declarative environment). Isn't this
  what you do in your formalization?}
\jimmy{TODO: a erasure function would make things more clear...}
\subsection{Soundness}

### 2019-02-11 sections/Overview.tex
0211 TODOs

@@

Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
the formalization harder since requires ``synchronizing'' things ...

@@

In recent work~\cite{itp2018}, we have shown how to mechanise a variant of DK's
subtyping algorithm and shown it correct with respect to DK's declarative
subtyping judgment. This approach overcomes the problem's in DK's formulation
by using a \emph{worklist}-based judgement of the form $\Gamma \vdash \Omega$
where $\Omega$ is a worklist, i.e. a sequence, of subtyping problems of the
form $A \leq B$.  The judgement is defined by case analysis on the first
element of $\Omega$ and recursively processes the worklist until it is empty.
Using both a single common ordered context $\Gamma$ and a worklist $\Omega$ greatly
simplifies propagating the instantiation of type variables that happen in one
subtyping problem to the remaining ones.

>>>>

%-------------------------------------------------------------------------------
\subsection{DK's Algorithm}
\jimmy{TODO: some of the text above about DK's algorithm may be moved here}

@@

\jimmy{TODO explain the duplicated context with some examples (their judgment form)}
Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
the formalization harder since requires ``synchronizing'' things ...

@@

In recent work~\cite{itp2018}, we have shown how to mechanise a variant of DK's
subtyping algorithm and shown it correct with respect to DK's declarative
subtyping judgment. This approach overcomes the problem's in DK's formulation
by using a \emph{worklist}-based judgement of the form $$\Gamma \vdash \Omega$$
\jimmy{emphasize the form of judgment}
where $\Omega$ is a worklist, i.e. a sequence, of subtyping problems of the
form $A \leq B$.  The judgement is defined by case analysis on the first
element of $\Omega$ and recursively processes the worklist until it is empty.
Using both a single common ordered context $\Gamma$ and a worklist $\Omega$ greatly
simplifies propagating the instantiation of type variables that happen in one
subtyping problem to the remaining ones.

### 2019-02-11 sections/conclusion.tex
0211 TODOs

\section{Conclusion and Future Work}

In this paper we have shown how to mechanise an algorithmic subtyping relation
for higher-order polymorphism, together with its proofs of soundness,
completeness and decidability, in the Abella proof assistant. 
In ongoing work we are extending our mechanisation with a bidirectional type
inference algorithm. The main difficulty there is communicating the
instantiations of existential variables from the subtyping algorithm to the
type inference. To make this possible we are exploring a continuation passing
style formulation, which generalises the worklist approach.
Another possible extension is to have the algorithm return an explicit witness
for the subtyping as part of type-directed elaboration into System F.

>>>>

\section{Conclusion}

This is the conclusion.

\jimmy{Discuss the efficient algorithm and its equivalence}
